---
title: "Chapter 12 Pictures vs Numbers"
output:
  word_document: default
  html_notebook: default
---

```{r}
Numberize <- function(inputVector){
  inputVector <- gsub(",","",inputVector)
  inputVector <- gsub(" ","",inputVector)
  return(as.numeric(inputVector))
}
```

#Takes string and removes commas and spaces, convets to number


#pull US2010 Census Data into a DataFrame

```{r}
readCensus <- function(){
testMunge <- read.csv(url("https://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv"))
testMunge <- testMunge[-1:-8,1:5]
testMunge <- testMunge[-52:-58,]
colnames(testMunge) <- c("stateName","april10census","april10base","july10pop","july11pop")
testMunge$stateName <- gsub("\\.","",testMunge$stateName)
testMunge$april10census <- Numberize(testMunge$april10census)
testMunge$april10base <- Numberize(testMunge$april10base)
testMunge$july10pop <- Numberize(testMunge$july10pop)
testMunge$july11pop <- Numberize(testMunge$july11pop)
rownames(testMunge) <- NULL
return(testMunge)}
```

Call the state data into R
```{r}
dfStates <- readCensus()
```

Create a historgram from Chapter 9
```{r}
hist(dfStates$july11pop)
```

Using base R graphics, we can also show a bar plot (a bar chart) with more detail
```{r}
barplot(dfStates$july11pop, names.arg = dfStates$stateName, las=2)
```

Create a function to check if a package is installed
```{r}
EnsurePackage <- function(x){
  x <- as.character(x)
  
  if (!require(x,character.only = TRUE)){
    install.packages(x, repos = "http://cran.r-project.org") 
    require(x,character.only = TRUE)}
}
```

Checking to see if ggplot2 is installed
```{r}
EnsurePackage("ggplot2")
```

Bulding a histogram in ggplot2

```{r}
g <- ggplot(dfStates, aes(x=july11pop))
g <- g + geom_histogram(binwidth = 5000000, color = "black", fill="white")
g <- g + ggtitle("states population histogram")
g
```

Creating a boxplot
```{r}
ggplot(dfStates, aes(x=july11pop,factor(0)))+ 
geom_boxplot()
```

graph population change
```{r}
dfStates$popChange <- dfStates$july11pop - dfStates$july10pop
dfStates$increasePop <- ifelse(dfStates$popChange > 0, "positive", "negative")
g <- ggplot(dfStates, aes(x=factor(increasePop), july11pop))
g <- g + geom_boxplot() + coord_flip()
g <- g + ggtitle ("Population grouped by positive or negative change")
g
```

Create a line Chart
```{r}
G <- ggplot(dfStates, aes(x=reorder(stateName, july11pop), y = july11pop, group=1))+geom_line()+theme(axis.text = element_text(angle = 90, hjust = 1))
G
```

Create a bar chart in ggplot
```{r}
G <- ggplot(dfStates, aes(x=reorder(stateName, july11pop), y = july11pop, group=1))+geom_col()+theme(axis.text = element_text(angle = 90, hjust = 1))
G
```

Population Percentage Change
```{r}
dfStates$percentChange <- dfStates$popChange/dfStates$july10pop*100
g <- ggplot(dfStates, aes(x=reorder(stateName, july11pop),y=july11pop,fill=percentChange))+geom_col()+theme(axis.text = element_text(angle = 90, hjust =1))
g
```

Create a scatter plot of the population change
```{r}
g <- ggplot(dfStates, aes(x=popChange, y=percentChange))+geom_point(aes(size=july11pop, color=july11pop))
g
```

Add text to show name ot the states
```{r}
g+geom_text(aes(label=stateName), size=4)
```

Adjust positions of the text
```{r}
g+geom_text(aes(label=stateName), size=4, hjust=1, vjust=-1)
```

```{r}
minPerChange <- 1
minPopChange <- 100000
dfStates$keystate <- dfStates$popChange > minPopChange & dfStates$percentChange > minPerChange
minLabel <- format(min(dfStates$july11pop), big.mark=",", trim=TRUE)
maxLabel <- format(max(dfStates$july11pop), big.mark=",", trim=TRUE)
medianLabel <- format(median(dfStates$july11pop, big.mark=",", trim=TRUE))
g <- ggplot(dfStates, aes(x=popChange, y=percentChange))+geom_point(aes(size=july11pop, color=july11pop, shape=keystate))
g <- g+geom_text(data=dfStates[dfStates$popChange>minPopChange & dfStates$percentChange>minPerChange,],aes(label=stateName, hjust=1, vjust=-1))
g+scale_color_continuous(name="Pop", breaks = with(dfStates, c(min(july11pop), median(july11pop), max(july11pop))), labels=c(minLabel,medianLabel, maxLabel), low="white",high="black")

```

