---
title: "Homework 1"
author: "Kyle Walter"
date: "10/6/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
IST687	– Intro Homework
Define	the	following	vectors,	which	represent	the	weight	and	height	of	people	on	a	
particular	team	(in	inches	and	pounds):
```{r}

			height	<- c(59,60,61,58,67,72,70)
			weight	<- c(150,140,180,220,160,140,130)
```

Define	a	variable:

```{r}
a	<- 150
```

Now	that	you	have	some	data	– explore	
Step	1:	Calculating	means
1) Compute,	using	R,	the	average	height	(called	mean	in	R)
```{r}
mean(height)
```

2) Compute,	using	R,	the	average	weight	(called	mean	in	R)
```{r}
mean(weight)
```

3) Calculate	the	length	of	the	vector	‘height’	and	‘weight’
```{r}
length(c(height,weight))
```

4) Calculate	the	sum	of	the	heights
```{r}
sum(height)
```

5) Compute	the	average of	both	height	and	weight,	by	dividing	the	sum (of	the	height	
or	the	width,	as	appropriate),	by	the	length	of	the	vector.	How	does	this	compare	to	
the	‘mean’	function?
```{r}
sum(height)/length(height)
sum(weight)/length(weight)
```
This produces the same value, however; there is more code to write rather that being able to quickly execute the code

Step	2:	Using	max/min	functions
6) Compute	the	max	height,	store	the	result	in	‘maxH’
```{r}
maxH<-max(height)
```

7) Compute	the	min	weight,	store	the	results	in	‘minW’
```{r}
minW<-min(weight)
```

Step	3:	Vector	Math
8) Create	a	new	vector,	which	is	the	weight	+	5	(every	person	gained	5	pounds)
```{r}
Covid5<-weight+5
```

9) Compute	the	weight/height	for	each	person,	using	the	new	weight	just created
```{r}
Covid5/height
```

Step	4:	Using	Conditional	if	statements
Hint:	In	R,	one	can	do:
if		(	100	<	150	)	"100	is	less	than	150"	else	"100	is	greater	than	150"
10)Write	the	R	code	to	test	if	max	height is	greater	than	60	(output	“yes”	or	“no”)
```{r}
if (maxH>60)"yes" else "no"
```

11)Write	the	R	code	to	if	min	weight is	greater	than	the	variable	‘a’	(output	“yes”	or	
“no”)
```{r}
if(minW>a) "yes" else "no"
```

