---
title: "Week 5 Homework"
author: "Kyle Walter"
date: "10/29/2020"
output:
  word_document: default
  pdf_document: default
---

IST687	– JSON	&	tapply Homework:	Accident	Analysis
Step	1:	Load	the	data
Read	in	the	following	JSON	dataset
http://data.maryland.gov/api/views/pdvh-tf2u/rows.json?accessType=DOWNLOAD

```{r}
library(RCurl)
library(jsonlite)
urltoRead <- "http://opendata.maryland.gov/api/views/pdvh-tf2u/rows.json?accessType=DOWNLOAD"
apiUrl <- getURL(urltoRead)
jData <- fromJSON(apiUrl)
mlanddf <- data.frame(jData[[2]])
```


Step	2:	Clean	the	data
After	you	load	the	data,	remove	the	first	8	columns,	and	then,	to	make	it	easier	to	work	
with,	name	the	rest	of	the	columns	as	follows:
Note,	not	surprisingly, it	is	in	JSON	format.		You	should	be	able	to	see	that	the	first	result	is	
the	metadata	(information	about	the	data)	and	the	second	is	the	actual	data.
namesOfColumns <-
c("CASE_NUMBER","BARRACK","ACC_DATE","ACC_TIME","ACC_TIME_CODE","DAY_OF_WE
EK","ROAD","INTERSECT_ROAD","DIST_FROM_INTERSECT","DIST_DIRECTION","CITY_NA
ME","COUNTY_CODE","COUNTY_NAME","VEHICLE_COUNT","PROP_DEST","INJURY","COLLI
SION_WITH_1","COLLISION_WITH_2")

```{r}
mlanddf <- mlanddf[,-1:-8]
namesOfColumns <-
c("CASE_NUMBER","BARRACK","ACC_DATE","ACC_TIME","ACC_TIME_CODE","DAY_OF_WEEK",
"ROAD","INTERSECT_ROAD","DIST_FROM_INTERSECT","DIST_DIRECTION","CITY_NAME",
"COUNTY_CODE","COUNTY_NAME","VEHICLE_COUNT","PROP_DEST","INJURY","COLLI
SION_WITH_1","COLLISION_WITH_2")
colnames(mlanddf) <- namesOfColumns
```


Step	3:	Understand	the	data	using	SQL	(via	SQLDF)
Answer	the	following	questions:
• How	many	accidents	happen	on	SUNDAY		
• How	many	accidents	had	injuries (might	need	to	remove	NAs	from	the	data)
• List	the	injuries	by	day

```{r}
library(sqldf)
sqldf("select count(DAY_OF_WEEK) as 'Accidents on Sundays' from mlanddf where TRIM(DAY_OF_WEEK) = 'SUNDAY'")
sqldf("select count(INJURY) as 'Accidents with injuries' from mlanddf where TRIM(INJURY) = 'YES'")
sqldf("select DAY_OF_WEEK, COUNT(INJURY) from mlanddf where INJURY = 'YES' GROUP BY DAY_OF_WEEK")

```


Step	4:	Understand	the	data	using	tapply
Answer	the	following	questions	(same	as	before)	– compare	results:
• How	many	accidents	happen	on	Sunday
• How	many	accidents	had	injuries (might	need	to	remove	NAs	from	the	data)
• List	the	injuries	by	day

```{r}
tapply(mlanddf$CASE_NUMBER, mlanddf$DAY_OF_WEEK, length)
tapply(mlanddf$CASE_NUMBER, mlanddf$INJURY, length)
tapply(mlanddf$CASE_NUMBER, list(mlanddf$DAY_OF_WEEK, mlanddf$INJURY), length)
```

