---
title: "Homework 3 - Cleaning/munging Dataframes"
author: "Kyle Walter"
date: "10/17/2020"
output:pdf_document: default
output:
  word_document: default
  pdf_document: default
---
Step	1:	Create	a	function	(named	readStates)	to	read	a	CSV	file	into	R
1. Note	that	you	are	to	read	a	URL,	not	a	file	local	to	your	computer.
2. The	file	is	a	dataset	on	state	populations	(within	the	United	States)

```{r}
readStates <- read.csv(url("http://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv"))
```

Step	2:	Clean	the	dataframe
3. Note	the	issues	that	need	to	be	fixed	(removing	columns,	removing	rows,	changing
column	names).
```{r}
str(readStates)
```


Need to remove the last 7 rows and first 8 rows from the data frame. Also need to remove columns X.4 through X.8 as they do not currently hold any values

Additionally need remove the . from the state name and convert column x through x.3 to numeric values from their character value.

4. Within	your	function,	make	sure	there	are	51	rows	(one	per	state	+	the	district	of
Columbia).	Make	sure	there	are	only	5	columns	with	the	columns	having	the
following	names	(stateName,	 base2010,	base2011,Jul2010, Jul2011).
5. Make	sure	the	last	four	columns	are	numbers	(i.e.	not	strings).

```{r}
readStates <- readStates[c(-1:-8,-60:-66),-6:-10]
```

```{r}
FreeNmbr <- function(inputvector){
  inputvector <- gsub(",","",inputvector)
  inputvector <- gsub(" ","",inputvector)
  return(as.numeric(inputvector))
}
```


```{r}
colnames(readStates) <- c("stateName", "base2010", "base2011", "Jul2010", "Jul2011")
```
```{r}
readStates$stateName <- gsub("\\.","",readStates$stateName)
readStates$base2010 <- FreeNmbr(readStates$base2010)
readStates$base2011 <- FreeNmbr(readStates$base2011)
readStates$Jul2010 <- FreeNmbr(readStates$Jul2010)
readStates$Jul2011 <- FreeNmbr(readStates$Jul2011)
```

Step	3:	Store	and	Explore	the	dataset
6. Store		the dataset	into	a	dataframe,	called	dfStates.
```{r}
dfStates <- readStates
```


7. Test	your	dataframe	by	calculating	the	mean	for	the	July2011	data,	by	doing:
mean(dfStates$Jul2011)
		 à you	should	get	an	answer	of		6,109,645
```{r}
mean(dfStates$Jul2011)
```
Step	4:	 Find	the	state	with	the	Highest	Population	
8. Based	on	the	July2011	data,	what	is	the	population	of	the	state	with	the	highest population?	What	is	the	name	of	that	state?

```{r}
dfStates[which.max(dfStates$Jul2011),1]
```

9. Sort	the	data,	in	increasing	order,	based	on	the	July2011	data.	

```{r}
dfStates <- dfStates[order(dfStates$Jul2011),]
```

Step	5:	 Explore	the	distribution	of	the	states

10. Write	a	function	that	takes	two	parameters.	The	first	is	a	vector	and	the	second	is	a	
number.
11. The	function	will	return	the	percentage	of	the	elements	within	the	vector	that	is	less	
than	the	same	(i.e.	the	cumulative	distribution	below	the	value	provided).

12. For	example,	if	the	vector	had	5	elements	(1,2,3,4,5),	with	2	being	the	number	
passed	into	the	function,	the	function	would	return	0.2	(since	20%	of	the	numbers	
were	below	2).



```{r}
library(scales)
cum.dist <- function(MyVector,MyNumber){
  p.below <- percent(length(which(MyVector<MyNumber))/length(MyVector))
  return(p.below)
}
```

Run a test to see if the functions work

```{r}
b <- c(1,2,3,4,5)
cum.dist(b,2)
```

13. Test	the	function	with	the	vector	‘dfStates$Jul2011Num’,	and	the	mean	of	
dfStates$Jul2011Num’.

```{r}
cum.dist(dfStates$Jul2011,mean(dfStates$Jul2011))
```
