---
title: "Homework 8"
author: "Kyle Walter"
date: "11/22/2020"
output:
  word_document: default
  pdf_document: default
---

IST687	– Making	Predictions
The	textbook’s	chapter	on	linear	models (“Line	Up,	Please”)	introduces	linear	predictive	
modeling	using	the	workhorse	tool	known	as	multiple	regression.	The	term	“multiple	
regression”	has	an	odd	history,	dating	back	to	an	early	scientific	observation	of	a	
phenomenon	called	“regression	to	the	mean.”	These	days,	multiple	regression	is	just	an	
interesting	name	for	using	a	simple	linear	modeling	technique	to	measuring	the	connection	
between	one	or	more	predictor	variables	and	an	outcome	variable In	this	exercise,	we	are	
going	to	use	an	open	data	set	to	explore	antelope	population.
This	is	the	first	exercise	of	the	semester	where	there	is	no	sample	R	code	to	help	you	along.	
Because	you	have	had	so	much	practice	with	R	by	now,	you	can	create	and/or	find	all	of	the	
code	you	need	to	accomplish	these	steps:
1. Read	in	data	from	the	following	URL:	
http://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/excel/mlr01.xls
This	URL	will	enable	you	to	download	the	dataset	into	excel.	
The	more	general	web	site	can	be	found	at:
http://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/frames/frame.html
If	you	view	this	in	a	spreadsheet,	you	will	find	that	four	columns	of	a	small	dataset.	
The	first	column	shows	the	number	of	fawn	in	a	given	spring	(fawn	are	baby	
Antelope).	The	second	column	shows	the	population	of	adult	antelope,	the	third	
shows	the	annual	precipitation	that	year,	and	finally,	the	last	column	shows	how	bad	
the	winter	was	during	that	year.

```{r}
library(readxl)
library(httr)

ULRtoRead <- "https://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/excel/mlr01.xls" #Captures the URL where the file is stored
GET(ULRtoRead, write_disk(data <- tempfile(fileext = ".xls"))) #Grabs the data from the website and stores it on the temporary drive --Credit-Stcack Overflow user LukeA for the design
Antelope <- read_xls(data) # Reads the data into the data frame from the temp file
colnames(Antelope) <- c("fawnPopulation","adultPopulation", "annualPercipitation","winterSeverity") #renames the columns to as described in the list above
print(Antelope) #shows the small data set to see that it meets spects described in the prompt -- it does
```


2. You	have	the	option	of	saving	the	file	save	this	file	to	your	computer	and	read	it	into	
R,	or reading the	data	directly	from	the	web	into	a	data	frame.

Read file Directly to data frame above - see 1

3. You	should	inspect	the	data	using	the	str()	command	to	make	sure	that	all	of	the	
cases	have	been	read	in	(n=8	years	of observations)	and	that	there	are	four
variables.	

```{r}
str(Antelope) #checks the data for the correct number over variables and size of the data frame - all good
```

4. Create	bivariate	plots	of	number	of	baby	fawns versus	adult	antelope	population,	
the	precipitation	that	year,	and	the	severity	of	the	winter.	Your	code	should	produce	
three	separate	plots.	Make	sure	the	Y-axis	and	X-axis	are	labeled.	Keeping	in	mind	
that	the	number	of	fawns is	the	outcome	(or	dependent)	variable,	which	axis	should	
it	go	on	in	your	plots?

```{r}
library(ggplot2)
#Graphs teh number of fawnsborn vs the Adult Population
AnteScatter <- ggplot(Antelope)+geom_point(aes(x=adultPopulation, y=fawnPopulation))+geom_smooth(method = lm, aes(x=adultPopulation, y=fawnPopulation), color="red")
#graphs the number of fawns to the amount of percipitation in the priot year.
rainScatter <- ggplot(Antelope)+geom_point(aes(x=annualPercipitation, y=fawnPopulation))+geom_smooth(method = lm, aes(x=annualPercipitation, y=fawnPopulation), color = "blue")
#graphs the #of Fawnsborn to the severity of the winter
winterScatter <- ggplot(Antelope)+geom_point(aes(x=winterSeverity, y=fawnPopulation))+geom_smooth(method = lm, aes(x=winterSeverity, y=fawnPopulation), color="green")

#the Three commands below call the graphs so they can be seen.
AnteScatter
rainScatter
winterScatter
```


5. Next,	create	three	regression	models	of	increasing	complexity	using	lm().	In	the	first	
model,	predict	the	number	of	fawns	from	the	severity	of	the	winter.	In	the	second	
model,	predict	the	number	of	fawns from	two	variables	(one	should	be	the	severity	
of	the	winter).	In	the	third	model	predict	the	number	of	fawns from	the	three	other	
variables.	Which	model	works	best?	

```{r}
model1 <- lm(fawnPopulation~winterSeverity, Antelope) # Creats linear regression predicting number of fawns born vs winter severity
summary(model1) # prints the results of Model1
model2 <- lm(fawnPopulation~winterSeverity+adultPopulation, Antelope) # Creats linear regresion predicting number of fawns born vs winter and adult population
summary(model2) #Prints the results of Model2
model3 <- lm(fawnPopulation~., Antelope) # creates linear model predicting fawns born vs all three dependent variables in the script
summary(model3) #prints results of Model3
```

Which	of	the	predictors	are	statistically	
significant	in	each	model? 

In the first model above, the prediction can only predict about 54% of the fawnsPopulation based on the severity of the winter. The Winter severity falls below the .05 P value meaning it is significant variable in the equation

In the 2nd model, I added in the adult population along with the winter severity, and 84% (Adjusted R Sqared value) of the fawn population can be predicted with this model THe Ault Population is a significant factor in the prediction with a p values less than .05 however; in this mix the severity of the winter has an increated p value above the .05 threshold and is no longer significant.

In the last model, the adjust R squared can predict 95.5% of the fawn population and all three factors are below the p value of .05 so they show as signficant. This model works the best of the three at predicting the number of fawnsborn.

If	you	wanted	to	create	the	most	parsimonious	model	(i.e.,	
the	one	that	did	the	best	job	with	the	fewest	predictors),	what	would	it	contain?

```{r}
summary(lm(fawnPopulation~adultPopulation, Antelope)) #Displays a linear model predicting fawns born vs the number of adults in the population.
```

I would utilize the equation to above as parsimonious model because I only need 1 variable the adult population of and can predict 88% of the time the fawn population. Both th Adult Population and Model have an overal p value less than .05 so it meets statistical significance.


Learning	Goals	for	this	activity:
A. Develop	skills	for	manipulating	and	transforming	data	that	contains	missing	values.
B. Understand	the	application	of	multiple	linear	regression	to	simple	situations	of	
predicting	one	numeric	variable	from	one	or	more	other	numeric	variables.	
C. Practice	plotting	skills.
D. Build	debugging	skills.
E. Increase	familiarity	with	bringing	external	data	sets	into	R.
F. Increase	familiarity	with	sources	of	advice	and	ideas	on	R	source	code.
Essential	Guide	for	All	IST687	Activities	(appears	at	the	end	of	all	activity	guides)
1. All	IST687	activities	work	on	what	some	people	call	a	“constructivist	learning”	
model.	By	developing	a	product	on	your	own,	testing	it	to	find	flaws,	improving	it,	
and	comparing	your	solution	to	the	solutions	of	other	people,	you	can	obtain	a	
deeper	understanding	of	a	problem,	the	tools	that	might	solve	that	problem,	and	a	
range	of	solutions	that	those	tools	may	facilitate.	The	constructivist	model	only	
works	to	the	extent	that	the	student/learner	has	the	drive	to	explore	a	problem,	be	
frustrated,	fail,	try	again,	possibly	fail	again,	and	finally	push	through	to	a	
satisfactory	level	of	understanding.
2. Each	IST687	activity	builds	on	skills	and	knowledge	developed	in	the	previous	
activities,	so	your	success	across	the	span	of	the	course	depends	at	each	stage	on	
your	investment	in	earlier	stages.	Take	the	time	to	experiment,	play, try	new	things,	
practice,	improve,	and	learn	as	much	as	possible.	These	investments	will	pay	off	
later.
3. Using	the	expertise	of	others,	the	Internet,	and	other	sources	of	information	is	not	
only	acceptable	- it	is	expected.	You	must	always,	always,	always give	credit	to	
your	sources.	For	example,	if	you	find	a	chunk	of	code	from	r-bloggers.com	that	
helps	you	with	developing	a	solution,	by	all	means	borrow	that	chunk	of	code,	but	
make	sure	to	use	a	comment	in	your	code	to	document	the	source	of	the	borrowed	
code	chunk.	The	discussion	boards	in	the	learning	management	system	have	been	
setup	to	encourage	appropriate	sharing	of	knowledge	and	wisdom	among	peers.	
Feel	free	to	ask	a	question	or	pose	a	solution	on	these	boards.
4. Building	on	the	previous	point,	when	submitting	code	as	your	solution	to	the	
activity,	the	comments	matter	at	least	as	much,	if	not	more	than	the	code	itself.	A	
good	rule	of	thumb	is	that	every	line	of	code	should	have	a	comment,	and	every	
meaningful	block	of	code	should	be	preceded	by	a	comment	block	that	is	just	about	
as	long	as	the	code	itself.	As	noted	above,	you	can	use	comments	to	give	proper	
credit	to	your	sources	and	you	can	use	comments	to	identify	your	submission	as	
your	own.		
5. Here’s	a	bonus	area	of	constructivist	learning	that	has	not	appeared	in	previous	
exercises	for	this	class:	Sometimes	the	building	process	reveals	unexpected	results	
that	are	themselves	very	informative	in	learning.	When	you	completed	the	exercise	
above,	what	did	you	find	that	was	unexpected?	What	did	you	do	about	trying	to	
understand	what	had	happened?	Did	you	do	further	exploration?	What	did	that	
further	exploration	reveal?

So I think that looking at winter severity on a graph was such that I would have dropped it as variable, it did not seem to play a signficant impact on the graph, but seeing that it did show signficance when run through the linear model was surprising. I guess if the mothers aren't making it through the winter the precived value will be less fawns born in the spring.