---
title: "Week 6 Homework"
author: "Kyle Walter"
date: "11/1/2020"
output: word_document
---

Step	1:	Load	the	data
We	will	use	the	airquality	data	set,	which you	should	already	have	as	part	of	your	R	
installation.
```{r}
df <- airquality
```

Step	2:	Clean	the	data
After	you	load	the	data,	there	will	be	some	NAs	in	the	data.	You	need	to	figure	out	what	to	
do	about	those	nasty	NAs.
```{r}
df <- na.omit(df)
colSums(is.na(df))
```


Step	3:	Understand	the	data	distribution
Create	the	following	visualizations using	ggplot:
• Histograms	for	each	of	the	variables	

```{r}
library(ggplot2)
ggplot(df, aes(Ozone))+geom_histogram(color="black", fill="gray")
ggplot(df, aes(Solar.R))+geom_histogram(color="black", fill="gray")
ggplot(df, aes(Wind))+geom_histogram(color="black", fill="gray")
ggplot(df, aes(x=Temp))+geom_histogram(color="black", fill="gray")
ggplot(df, aes(x=Month))+geom_histogram(color="black", fill="gray")
ggplot(df, aes(x=Day))+geom_histogram(color="black", fill="gray")


```


• Boxplot	for	Ozone
```{r}
ggplot(df, aes(as.factor(Ozone),Month))+geom_boxplot()
```

• Boxplot for	wind	values	(round	the	wind	to	get	a	good	number	of	“buckets”)
```{r}
ggplot(df, aes(as.factor(Wind),Month))+geom_boxplot()
```

Step	4:	Explore	how	the	data	changes	over	time
First,	make	sure	to	create	appropriate	dates	(this	data	was	from	1973).	Then	create	line	
charts for	ozone,	temp,	wind	and	solar.R (one	line	chart	for	each,	and	then	one	chart	with	4	
lines,	each	having	a	different	color).	Create	these	visualizations	using	ggplot.
Note	that	for	the	chart	with	4	lines,	you	need	to	think	about	how	to	effectively	use	the	yaxis.

```{r}
df$Date <- paste(1973, df$Month, df$Day, sep = "-")
df$Date <- as.Date(df$Date, tryFormats=c("%Y-%m-%d", "%Y/%m/%d"))
Variables <- c(df$Ozone,df$Temp, df$Wind,df$Solar.R)
Variables_Name <- rep(c("Ozone","Temp",
                      "Wind","Solar.R"),each = 111)
Days <- rep(df$Date,4)
df2 <- data.frame(Variables,Variables_Name, Days)
ggplot(df, aes(x=Date, y=Ozone))+geom_line()+geom_point()
ggplot(df, aes(x=Date, y=Solar.R))+geom_line()+geom_point()
ggplot(df, aes(x=Date, y=Temp))+geom_line()+geom_point()
ggplot(df, aes(x=Date, y=Wind))+geom_line()+geom_point()
g <- ggplot(df2, aes(x=Days, y=Variables, Group=Variables_Name, color=Variables_Name))+geom_line()+geom_point()
g
```


Step	5:	Look	at	all	the data	via	a	Heatmap
Create a	heatmap,	with	each	day	along	the	x-axis	and	ozone,	temp,	wind	and	solar.r	along	
the	y-axis,	and	days	as	rows	along	the	y-axis.	 Great	the	heatmap	using	geom_tile (this	
defines	the	ggplot geometry	to	be	‘tiles’	as	opposed	to	‘lines’	and	the	other	geometry	we	
have	previously	used).
Note	that	you	need	to	figure	out	how	to	show	the	relative	change	equally	across	all	the	
variables.

```{r}
ggplot(df2, aes(x=as.factor(Days), y=as.factor(Variables_Name), fill = Variables))+geom_tile()+xlab("Date(in1973)")+ylab("Variables")+labs(fill="Value")
```


Step	6:	Look	at	all	the	data	via	a	scatter	chart
Create	a	scatter	chart (using	ggplot	geom_point),	with	the	x-axis	representing	the	wind,	the	
y-axis	representing	the	temperature,	the	size	of	each	dot	representing	the	ozone	and	the	
color	representing	the	solar.R

```{r}
ggplot(df, aes(x=Wind, y=Temp, color = Solar.R, size = Ozone))+geom_point()
```


Step	7:	Final	Analysis
• Do	you	see	any	patterns	after	exploring	the	data?		

Solar Radiation and Ozone definitely seem to have a correlation. When one of the two variable is high so is the other. And this is easily seeon the line graph and reinforced on the scatter plot. Warmer temperatues also have a correlation to slower wind speed. There also seems to be relation between warmer temperatures and and Higher lvels of Ozone and Solar Radiation. Though looking at the lin chart Solar R leads Ozone which leads tempture.

• Which image was the most helpful
 I think for the data set the line graph and scatter plot working together tell a larger story. There we items I noticed on the line graph, but upon revisiting after seeing the scatter plot, I found more useful information.







