---
title: "Week 4 Homework"
author: "Kyle Walter"
date: "10/29/2020"
output:
  word_document: default
  pdf_document: default
---

IST687	– Samples	HW
Let’s	continue	our	exploration	of	sampling.
Step	1:	Write	a	summarizing	function	to	understand	the	distribution	of	a	vector
1. The	function,	call	it	‘printVecInfo’ should	take	a	vector	as	input
2. The	function	should	print	the	following	information:
a. Mean
b. Median
c. Min	&	max
d. Standard	deviation
e. Quantiles	(at	0.05	and	0.95)
f. Skewness

```{r}
library(moments)
printVectorInfo <- function(myVector) {
  meanVector <- mean(myVector)
  medianVector <- median(myVector)
  minVector <- min(myVector)
  maxVector <- max(myVector)
  sdVector <- sd(myVector)
  qVector <- quantile(myVector, probs = c(.05, .95))
  skewnessVector <- skewness(myVector)
  cat('Mean: ',meanVector,'\n')
  cat('Median: ',medianVector, '\n')
  cat('Min: ',minVector,' ')
  cat('Max: ',maxVector,'\n')
  cat('Std: ',sdVector,'\n')
  cat('Quantile (.05 -- .95): ',qVector,'\n')
  cat('Skewness: ',skewnessVector,'\n\n')
}
```

Note	for	skewness,	you	can	use	the	function	in	the	‘moments’	library.00
3. Test	the	function	with	a	vector	that	has	(1,2,3,4,5,6,7,8,9,10,50).	You	should	see	
something	such	as:
[1]	"mean:	9.54545454545454"
[1]	"median:	6"
[1]	"min:	1		max:	50"
[1]	"sd:	13.7212509368762"
[1]	"quantile	(0.05	- 0.95):	1.5	-- 30"
[1]	"skewness:	2.62039633563579"
```{r}
testVector <- c(1,2,3,4,5,6,7,8,9,10,50)
printVectorInfo(testVector)
```


Step	2:	Creating	Samples	in	a	Jar
4. Create	a	variable	‘jar’	that	has	50	red	and	50	blue	marbles
(hint:	the	jar	can	have	strings	as	objects,	with	some	of	the	strings	being	‘red’	and	
some	of	the	strings	being	‘blue’

```{r}
jar <- rep(c("red","blue"),50)
```


5. Confirm	there	are	50	reds	by	summing	the	samples	that	are	red

```{r}
sum(jar=="red")
```


6. Sample	10	‘marbles’	(really	strings)	from	the	jar.	How	many	are	red?	What	was	the	
percentage	of	red	marbles?

```{r}
library(scales)
MarSample <- sample(jar, 10, replace = TRUE)
percent(sum(MarSample=="red")/10)
```


7. Do	the	sampling	20	times,	using	the	‘replicate’	command.	This	should	generate	a	list	
of	20	numbers.	Each	number	is	the	mean	of	how	many	reds	there	were	in	10	
samples.	Use	your	printVecInfo to	see	information	of	the	samples.	Also	generate	a	
histogram	of	the	samples.

```{r}
MarRepSample <- replicate(20, length(which((sample(jar,10, replace = TRUE ))=="red")))
mean(MarRepSample)
printVectorInfo(MarRepSample)
hist(MarRepSample)
```


8. Repeat	#7,	but	this	time,	sample	the	jar	100	times.	You	should	get	20 numbers,	this	
time	each	number	represents	the	mean	of	how	many	reds	there	were	in	the	100
samples.	Use	your	printVecInfo	to	see	information	of	the	samples.	Also	generate	a	
histogram	of	the	samples.

```{r}
MarRepSample2<- replicate(20, length(which((sample(jar,100, replace = TRUE))=="red")))
mean(MarRepSample2)
printVectorInfo(MarRepSample2)
hist(MarRepSample2)
```

9. Repeat	#8,	but	this	time,	replicate	the	sampling	100 times.	You	should	get	100	
numbers,	this	time	each	number	represents	the	mean	of	how	many	reds	there	were	
in	the	100	samples.	Use	your	printVecInfo	to	see	information	of	the	samples.	Also	
generate	a	histogram	of	the	samples.

```{r}
MarRepSample3<- replicate(100, length(which((sample(jar,100, replace = TRUE))=="red")))
mean(MarRepSample3)
printVectorInfo(MarRepSample3)
hist(MarRepSample3)
```


Step	3:	Explore	the	airquality	dataset
10. Store	the	‘airquality’	dataset	into	a	temporary	variable

```{r}
AirQ <- airquality
```


11. Clean	the	dataset	(i.e.	remove	the	NAs)
```{r}
summary(AirQ)
AirQ <- AirQ[rowSums(is.na(AirQ))<=0,]
summary(AirQ)
```


12. Explore	Ozone,	Wind	and	Temp	by	doing	a	‘printVecInfo’	on	each	as	well	as	
generating	a	histogram	for	each

```{r}
printVectorInfo(AirQ$Ozone)
printVectorInfo(AirQ$Wind)
printVectorInfo(AirQ$Temp)
hist(AirQ$Ozone)
hist(AirQ$Wind)
hist(AirQ$Temp)
```

